---
title: "tidydice"
author: "Roland Krasser"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{explore}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

A basic understanding of probability and statistics is crucial for data understanding. A great way to teach is to start with an experiment, like rolling a dice or flipping a coin.

This package simulates rolling a dice and flipping a coin generating tidy data.

tidydice package on Github: [https://github.com/rolkra/tidydice](https://github.com/rolkra/tidydice)

### Setup

As the tidydice-functions fits well into the tidyverse, we load the dplyr-package. For quick visualisations we use the explore-package.

```{r message=FALSE, warning=FALSE}
library(tidydice)
library(dplyr)
library(explore)
```

### Roll a dice

#### Roll a dice once

```{R echo=TRUE}
set.seed(123)
roll_dice()
```

The result is 2, so it is no success (as only 6 is predefined as success)

#### Roll a dice 4 times

```{R echo=TRUE}
set.seed(123)
roll_dice(times = 4)
```

We got 1 success

#### Define rounds

```{R echo=TRUE}
set.seed(123)
roll_dice(times = 4, rounds = 2)
```

In the first round we got 1 success, in the secound round 2 success.

#### Use agg

A convenient way to aggregate the result, is to use the agg parameter

```{R echo=TRUE}
set.seed(123)
roll_dice(times = 4, rounds = 2, agg = TRUE)
```

#### Visualise result

You can use any package/tool you like to visualise the result. In this example we use the explore-package.

```{R echo=TRUE}
set.seed(123)
roll_dice(times = 100) %>% 
  explore(result, title = "Rolling a dice 100x")
```

In 15% of the cases we got a six. This is close to the expected value of 100/6 = 16.67%

If we increase the times parameter to 10000, the results are more balanced.

```{R echo=TRUE}
set.seed(123)
roll_dice(times = 10000) %>% 
  explore(result, title = "Rolling a dice 100x")
```

#### Visualise success

If we repeat the experiment (rolling a dice 100x) with rounds = 100, we get the distribution with a peak at about 17 (16.67 is the expected value)

```{R echo=TRUE}
set.seed(123)
roll_dice(times = 100, rounds = 100, agg = TRUE) %>% 
  explore(success, 
          title = "Rolling a dice 100x",
          auto_scale = FALSE)
```

If we increase rounds from 100 to 10000 we get a more symmetric shape. We see that success below 5 and success above 30 are very unlikely.

```{R echo=TRUE}
set.seed(123)
roll_dice(times = 100, rounds = 10000, agg = TRUE) %>% 
  explore(success, 
          title = "Rolling a dice 100x",
          auto_scale = FALSE)
```

### Combine experiments

Let's add an experiment, where you have 10 extra dice. The shape changes.

```{R echo=TRUE}
set.seed(123)
roll_dice(times = 100, rounds = 10000, agg = TRUE) %>% 
  roll_dice(times = 110, rounds = 10000, agg = TRUE) %>% 
  explore(success, 
          target = experiment,
          title = "Rolling a dice 100x / 110x",
          auto_scale = FALSE)
```

You can add as many experiments as you like (as long they generate the same data structure)

Adding an experiment with times = 150 will generate a smaller but wider shape.

```{R echo=TRUE}
set.seed(123)
roll_dice(times = 100, rounds = 10000, agg = TRUE) %>% 
  roll_dice(times = 110, rounds = 10000, agg = TRUE) %>% 
  roll_dice(times = 150, rounds = 10000, agg = TRUE) %>% 
  explore(success, 
          target = experiment,
          title = "Rolling a dice 100x / 110x / 150x",
          auto_scale = FALSE)
```
