---
title: "tidydice"
author: "Roland Krasser"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tidydice}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Introduction

A basic understanding of probability and statistics is crucial for data understanding and discovery of meaningful patterns. A great way to teach probability and statistics is to start with an experiment, like rolling a dice or flipping a coin.

This package simulates rolling a dice and flipping a coin. Each experiment generates a tibble. Dice rolls and coin flips are simulated using sample(). The properties of the dice can be changed, like the number of sides. A coin flip is simulated using a two sided dice. Experiments can be combined with the pipe-operator. 

tidydice package on Github: [https://github.com/rolkra/tidydice](https://github.com/rolkra/tidydice)

### Setup

As the tidydice-functions fits well into the tidyverse, we load the dplyr-package. For quick visualisations we use the explore-package. To create more flexible graphics, use ggplot2.

```{r message=FALSE, warning=FALSE}
library(tidydice)
library(dplyr)
library(explore)
```

### Roll a dice

#### Example

6 dice are rolled 3 times using roll_dice(). The result of the dice-experiment is visualised using plot_dice().

```{r fig.height=3, fig.width=6}
set.seed(123)
roll_dice(times = 6, rounds = 3) %>% 
  plot_dice(fill_success = "lightgrey")
```

#### Roll a dice once

The output of roll_dice() is a tibble. Each row represents a dice roll. Without parameters, a dice is rolled once. You can use plot_dice() to visualise the result.

```{R echo=TRUE}
set.seed(123)
roll_dice()
```

```{R echo=TRUE, fig.width=6, fig.height=1}
set.seed(123)
roll_dice() %>% plot_dice()
```

Success is defined as result = 6 (as default), while result = 1..5 is not a success. In this case the result is 2, so it is no success.

#### Define success

If we would define result = 2 and result = 6 as success, it would be treated as success.

```{R echo=TRUE}
set.seed(123)
roll_dice(success = c(2,6))
```

#### Unfair dice

As default, the dice is fair. So every result (0..6) has the same probability. If you want, you can change this.

```{R echo=TRUE}
set.seed(123)
roll_dice(prob = c(0,0,0,0,0,1))
```

In this case we created a dice that always gets result = 6 (with 100% probability)

#### Untypically dice

As default the dice has 6 sides. If you want you can change this. Here we use a dice with 12 sides. result now can have a value between 1 and 12. But result = 6 is still the default success.

```{R echo=TRUE}
set.seed(123)
roll_dice(sides = 12)
```

#### Roll a dice 4 times

```{R echo=TRUE}
set.seed(123)
roll_dice(times = 4)
```

```{R echo=TRUE, fig.width=6, fig.height=1}
set.seed(123)
roll_dice(times = 4) %>% plot_dice()
```

We get 1 success

#### Define rounds

```{R echo=TRUE}
set.seed(123)
roll_dice(times = 4, rounds = 2)
```

```{R echo=TRUE, fig.width=6, fig.height=2}
set.seed(123)
roll_dice(times = 4, rounds = 2) %>% plot_dice()
```

Rolling the dice 4 times is repeated. In the first round we got 1 success, in the secound round 2 success.

#### Use agg

A convenient way to aggregate the result, is to use the agg parameter. Now we get one line per round.

```{R echo=TRUE}
set.seed(123)
roll_dice(times = 4, rounds = 2, agg = TRUE)
```

You can aggregate by hand too using dplyr.

```{R echo=TRUE}
set.seed(123)
roll_dice(times = 4, rounds = 2) %>% 
  group_by(experiment, round) %>% 
  summarise(times = n(),
            success = sum(success))
```


#### Visualise result

You can use any package/tool you like to visualise the result. In this example we use the explore-package.

```{R echo=TRUE, fig.height=4, fig.width=7}
set.seed(123)
roll_dice(times = 100) %>% 
  explore(result, title = "Rolling a dice 100x")
```

In 15% of the cases we got a six. This is close to the expected value of 100/6 = 16.67%

If we increase the times parameter to 10000, the results are more balanced.

```{R echo=TRUE, fig.height=4, fig.width=7}
set.seed(123)
roll_dice(times = 10000) %>% 
  explore(result, title = "Rolling a dice 10000x")
```

#### Visualise success

If we repeat the experiment rolling a dice 100x with rounds = 100, we get the distribution with a peak at about 17 (16.67 is the expected value)

```{R echo=TRUE, fig.height=4, fig.width=7}
set.seed(123)
roll_dice(times = 100, rounds = 100, agg = TRUE) %>% 
  explore(success, 
          title = "Rolling 100 dice 100x",
          auto_scale = FALSE)
```

If we increase rounds from 100 to 10000 we get a more symmetric shape. We see that success below 5 and success above 30 are very unlikely.

```{R echo=TRUE, fig.height=4, fig.width=7}
set.seed(123)
roll_dice(times = 100, rounds = 10000, agg = TRUE) %>% 
  explore(success, 
          title = "Rolling 100 dice 10000x",
          auto_scale = FALSE)
```

This shape is already very close to the binomial distribution

```{R echo=TRUE, fig.height=4, fig.width=7}
binom_dice(times = 100) %>% 
  plot_binom(title = "Binomial distribution, rolling 100 dice")
```

```{R echo=TRUE}
set.seed(123)
roll_dice(times = 100, rounds = 10000, agg = TRUE) %>% 
  mutate(check = ifelse(success < 5 | success > 30, 1, 0)) %>% 
  count(check)
```

In only 4 of 10000 (0.04%) cases success is below 5 or above 30. So the probability to get this result is very low.

We can check that with the binomial distribution too:

```{R echo=TRUE}
binom_dice(times = 100) %>% 
  filter(success < 5 | success > 30)
```

```{R echo=TRUE}
binom_dice(times = 100) %>% 
  filter(success < 5 | success > 30) %>% 
  summarise(check_pct = sum(pct))
```

The probability to get this result is 0.04% (based on the binomial distribution).

#### Combine experiments

Let's add an experiment, where you have 10 extra dice. The shape of the distribution changes.

```{R echo=TRUE, fig.height=4, fig.width=7}
set.seed(123)
roll_dice(times = 100, rounds = 10000, agg = TRUE) %>% 
  roll_dice(times = 110, rounds = 10000, agg = TRUE) %>% 
  explore(success, 
          target = experiment,
          title = "Rolling a dice 100/110x",
          auto_scale = FALSE)
```

You can add as many experiments as you like (as long they generate the same data structure)

Adding an experiment with times = 150 will generate a smaller but wider shape.

```{R echo=TRUE, fig.height=4, fig.width=7}
set.seed(123)
roll_dice(times = 100, rounds = 10000, agg = TRUE) %>% 
  roll_dice(times = 110, rounds = 10000, agg = TRUE) %>% 
  roll_dice(times = 150, rounds = 10000, agg = TRUE) %>% 
  explore(success, 
          target = experiment,
          title = "Rolling a dice 100/110/150x",
          auto_scale = FALSE)
```

#### Binomial distribution

Rolling a dice 100x, a result between 10 and 23 has a probability of over 94%

```{R echo=TRUE, fig.height=4, fig.width=7}
binom_dice(times = 100) %>% 
  plot_binom(highlight = c(10:23))
```

### Flip a coin

Internally the package handles coins as dice with only two sides. Success is defined as result = 2 (as default), while result = 1 is not a success.

#### Flip a coin 10x

```{R echo=TRUE}
set.seed(123)
flip_coin(times = 10)
```

In this case the result are 6x 2 and 4x 1. We can use the describe() function of the explore-package to get a good overview.

```{R echo=TRUE}
set.seed(123)
flip_coin(times = 10) %>% 
  describe(success)
```

Or just use the agg-parameter

```{R echo=TRUE}
set.seed(123)
flip_coin(times = 10, agg = TRUE)
```

#### Define rounds

The parameter rounds can be used like in roll_dice().

```{R echo=TRUE}
set.seed(123)
flip_coin(times = 10, rounds = 4, agg = TRUE)
```

#### Combine experiments

```{R echo=TRUE}
set.seed(123)
flip_coin(times = 10, agg = TRUE) %>% 
  flip_coin(times = 15, agg = TRUE)
```

#### Binomial Distribution

```{R echo=TRUE}
binom_coin(times = 10) 
```

```{R echo=TRUE}
binom_coin(times = 10) %>% 
  plot_binom(title = "Binomial distribution,\n10 coin flips")
```

### Dice design

#### Default settings

```{r fig.height=1, fig.width=6}
set.seed(123)
roll_dice(times = 6) %>% 
  plot_dice()
```

#### Change color

```{r fig.height=1, fig.width=6}
set.seed(123)
roll_dice(times = 6) %>% 
  plot_dice(fill = "black", line_color = "white", point_color = "white")
```

```{r fig.height=1, fig.width=6}
set.seed(123)
roll_dice(times = 6) %>% 
  plot_dice(fill = "lightblue", fill_success = "gold")
```

```{r fig.height=1, fig.width=6}
set.seed(123)
roll_dice(times = 6) %>% 
  plot_dice(fill = "darkgrey", 
            fill_success = "darkblue",
            line_color = "white",
            point_color = "white")
```

#### Dice type

```{r fig.height=1, fig.width=6}
set.seed(123)
roll_dice(times = 6) %>% 
  plot_dice(detailed = TRUE)
```

```{r fig.height=1, fig.width=6}
set.seed(123)
roll_dice(times = 6) %>% 
  plot_dice(detailed = FALSE)
```

#### Limits

plot_dice() is limited to 1 experiment with max. 10 times x 10 rounds.

```{r fig.height=6, fig.width=6}
set.seed(123)
roll_dice(times = 10, rounds = 10) %>% 
  plot_dice(detailed = FALSE, fill_success = "gold")
```

#### Force

You can force a result using force_dice() and force_coin().

```{r fig.height=1, fig.width=6}
force_dice(1:6) %>% 
  plot_dice()
```

```{r fig.height=1, fig.width=6}
force_dice(rep(6, times = 6)) %>% 
  plot_dice()
```

We can combine two foreced dice rolling using the pipe operator and the parameter round.

```{r fig.height=1, fig.width=6}
force_dice(rep(5, times = 3), round = 1) %>% 
  force_dice(rep(6, times = 3), round = 2)
```

```{r fig.height=1, fig.width=6}
set.seed(123)
force_dice(rep(6, times = 3)) %>% 
  roll_dice(times = 3)
```

In the first experiment we get 3 times a 6 (forced), but in the second experiment none.

### Using Dice Formula

If you want to do more complex dice rolls, use ```roll_dice_formula()```

```{r}
# roll 1 dice with 6 sides
roll_dice_formula(dice_formula = "1d6", seed = 123)
```

```{r}
roll_dice_formula(
  dice_formula = "4d6", # 4 dice with 6 sides
  success = 15:24,      # success is defined as sum between 15 and 24
  seed = 123            # random seed to make it reproducible
)
```

```{r}
# roll 4 dice with 6 sides
roll_dice_formula(
  dice_formula = "4d6", # 4 dice with 6 sides
  rounds = 10,          # repeat 10 times
  success = 15:24,      # success is defined as sum between 15 and 24
  seed = 123            # random seed to make it reproducible
)
```

```{r}
roll_dice_formula(
  dice_formula = "4d6e3", # 4 dice with 6 sides, explode on a 3
  rounds = 5,             # repeat 5 times
  success = 15:24,        # success is defined as sum between 15 and 24
  seed = 123              # random seed to make it reproducible
)
```

```{r}
roll_dice_formula(
  dice_formula = "4d6+1d10", # 4 dice with 6 sides + 1 dice with 10 sides
  rounds = 1000) %>%         # repeat 1000 times
  explore_bar(result, numeric = TRUE)  # visualise result
```

Other examples for dice_formula:

- ```1d6``` = roll one 6-sided dice
- ```1d8``` = roll one 8-sided dice
- ```1d12``` = roll one 12-sided dice
- ```2d6``` = roll two 6-sided dice
- ```1d6e6``` = roll one 6-sided dice, explode dice on a 6
- ```3d6kh2``` = roll three 6-sided dice, keep top 2 rolls
- ```3d6kl2``` = roll three 6-sided dice, keep lowest 2 rolls
- ```4d6kh3e6``` = roll four 6-sided dice, keep top 3 rolls, but explode on a 6
- ```1d20+4``` = roll one 20-sided dice, and add 4
- ```1d4+1d6``` = roll one 4-sided dice and one 6-sided dice, and sum the results